### 2019-02-27 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Finishing Chapter:4

We consider two hypotheses $\cH_{1}$ and $\cH_{2}$ defined as,
\begin{itemize}
	\item $\cH_{1}$:
    The probability $p_{c1}$ is composed by the sum of two probabilities: i)
    the probability of colliding consecutively twice, \ie, the devices that collide at a given time slot and collide again when retransmitting their packets,
    and ii) the probability of collision related with devices that attempt to transmit in this channel for first time.
    In addition, we suppose that the number of devices that attempt to retransmit is small in comparison to the total number of devices sharing the same channel.

	\item $\cH_{2}$:
	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
\end{itemize}

@@

@@

As explained above, if an IoT device is experiencing a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
We denote $p_{ca}$ the probability to have a collision with a packet involved in the previous collision.
Under assumption $\cH_{1}$, the probability that the same device's packet is experiencing again a collision at the second time slot is
\begin{equation}\label{eq:43:decomppc1}
	p_{c1} = p_{ca}+\left(1-p_{ca} \right)p_c.
\end{equation}

If the device had a collision, we consider $p_{bp}(n)$ the probability that it had a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in the previous collision chose the same back-off interval.
Hence, under hypothesis $\cH_{2}$, we can relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
%
\begin{equation}\label{eq:43:sumpca}
	p_{ca} = \sum_{n=1}^{N-1} p_{bp}(n).
\end{equation}

Therefore, the expression of $p_{ca}$ is
\begin{equation}\label{eq:43:sumpc2}
	\frac{1}{p_c} \sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right].
\end{equation}

Once again under $\cH_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first terms of the sum in \eqref{eq:43:sumpc2} are predominant. Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$. Thus it gives,
\begin{align}\label{eq:43:sumpca2}
	p_{ca} = & 1- \frac{1}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left( 1-\frac{1}{m}\right)^n,\nonumber\\
	\simeq & 1- \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N-1}{N-1 \choose k} x^n \left( 1-\frac{1}{m}\right)^n.
\end{align}

We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca2}, and so the expression of $p_{ca}$ becomes
\begin{align}\label{eq:43:pca}
	p_{ca} \simeq \frac{1}{p_c} - \left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
\end{align}

>>>>

We consider two hypotheses $\mathcal{H}_{1}$ and $\mathcal{H}_{2}$ defined as,
\begin{itemize}
	\item $\mathcal{H}_{1}$:
    The probability $p_{c1}$, is composed by the sum of two probabilities: i)
    the probability of colliding consecutively twice, \ie, the devices that collide at a given time slot and collide again when retransmitting their packets,
    and ii) the probability of collision among devices that did not collide in the same previous collision. Moreover, we suppose that the number of devices involved in a collision is small in comparison to the total number of devices.
	\item $\mathcal{H}_{2}$:
	The total number of the back-off stages at time $t$ is constant, and it is assumed to be large enough to consider that no device will ever be in the last failure state (this case is the one on the right side in Figure~\ref{fig:43:Markov_model}), after $M$ successive failed retransmissions.
\end{itemize}

@@

%{\color{red}Considering one device and a channel,
%we denote $x_t^i$ the packet transmission probability for the $i+1$ time in a given time slot $t$ (with $i\in \llbracket 0, M-1 \rrbracket$).}
%
%{\color{red}Considering one device and a channel, we denote $x_t^i$ the transmission probability for the $i+1$ time, with $i\in \llbracket 0, M-1 \rrbracket$, in a given time slot $t$}

@@

As explained above, if an IoT device experiences a collision at the first transmission, it proceeds for the retransmission of its packet after a random back-off interval.
We denote $p_{ca}$ the probability to have a collision with a packet involved in the previous collision.
Under the $\mathcal{H}_{1}$ assumption, the number of packets involved in the same previous collision remains very small in comparison to the total number of devices that may transmit during this time. In other words, this collision probability does not depend on previous retransmissions and is equal to $p_c$.
So, the probability that the same device's packet experiences again a collision at the second time slot is
%
\begin{equation}\label{eq:43:decomppc1}
	p_{c1} = p_{ca}+\left(1-p_{ca} \right)p_c.
\end{equation}

If the device has a collision at the first attempt, we consider $p_{bp}(n)$ the probability that it has a collision with \emph{exactly} $n$ packets (for any $1\leq n \leq N-1$), and that \emph{at least one} of the $n$ devices involved in this first collision chooses the same back-off interval,
%
\begin{equation}
    p_{bp}(n) = {N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right].
\end{equation}

Besides, $p_{ca}$ is the conditional probability of collision with a packet sent by a device involved in the previous collision given that the packet experienced collision at its first transmission.
Hence, under hypothesis $\mathcal{H}_{2}$, we can use Bayes theorem and the law of total probability to relate $p_{ca}$ with $p_{bp}(n)$, and the different probabilities that a device experienced a collision during the first slot and has the same back-off interval for its retransmission is,
%{\color{red}Hence, under hypothesis $\mathcal{H}_{2}$,
%the collision probability with the same packets after a retransmission, $p_{ca}$, is given  the sum of collision probabilities with $n$ devices choose the same back-off interval for their retransmissions, }
%
\begin{equation}\label{eq:43:sumpca}
	p_{ca} = \frac{1}{p_c}\sum_{n=1}^{N-1} p_{bp}(n).
\end{equation}

% \begin{figure}[htp!]  % [htbp]
% 	\centering
% 	% \includegraphics[width=1.00\linewidth]{Approximation_m10.eps}
% 	\includegraphics[width=1.00\linewidth]{Approximation_m10-eps-converted-to.pdf}
% 	\caption{Our proposed approximation for the probability of collision at the second transmission. It is more precise for smaller values of $N$.}
% 	\label{fig:43:Approximation_m10}
% \end{figure}

Therefore, the expression of $p_{ca}$ is
\begin{align}\label{eq:43:sumpc2}
	& \frac{1}{p_c} \sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left[1-\left( 1-\frac{1}{m}\right)^n \right]\nonumber \\
	& = 1- \frac{1}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left(1-x\right)^{N-1-n}\left( 1-\frac{1}{m}\right)^n.
\end{align}

Once again under $\mathcal{H}_{1}$, assuming that the number of devices involved in the first collision is small compared to $N-1$, the first $N_0 \ll N-1$ terms of the sum in \eqref{eq:43:sumpc2} are predominant. We derive,
%
\begin{equation}\label{eq:43:sumpca2}
	p_{ca} \simeq  1- \frac{1}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n\left(1-x\right)^{N-1-n} \left( 1-\frac{1}{m}\right)^n.
\end{equation}

Moreover, for these terms, $n$ is small compared to $N-1$, and so $N-1-n$ can be approximated to $N-1$. Thus it gives,
%
\begin{equation}\label{eq:43:sumpca3}
	p_{ca} \simeq  1- \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N_0}{N-1 \choose n} x^n \left( 1-\frac{1}{m}\right)^n.
\end{equation}

Assuming $\mathcal{H}_{1}$ amounts to consider that $x\ll 1$. As a consequence, the sum in equation \eqref{eq:43:sumpca3} can be supplemented by negligible terms,

\begin{equation}\label{eq:43:sumpca4}
	p_{ca} \simeq  1 - \frac{\left(1-x\right)^{N-1}}{p_c}\sum_{n=1}^{N-1}{N-1 \choose n} x^n \left( 1-\frac{1}{m}\right)^n.
\end{equation}

We use the binomial theorem to compute the sum in \eqref{eq:43:sumpca4}, and we rewrite the expression of $p_{ca}$ as
\begin{multline}\label{eq:43:pca}
	p_{ca} \simeq 1 -\\
	\left(\frac{1}{p_c}-1\right)\left[ 1+\left(1-\left(1-p_c\right)^{\frac{1}{N-1}}\right)\left(1-\frac{1}{m}\right)\right]^{N-1}.
\end{multline}

### 2019-02-27 2-Chapters/4-Chapter/chapter4.tex
Finishing Chapter:4

% ----------------------------------------------------------------------
\chapter{More Efficient Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc
% ----------------------------------------------------------------------

@@

More precisely, on one hand, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, used by static devices.
Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their communication messages were successfully received.
We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time. If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
we show in simulations that in the case of non-uniform repartition of static devices in the channels, the dynamic devices can improve a lot the network efficiency.
%
On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real radio hardware.

On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending the first one, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
Interestingly, the simpler heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.

@@

This Chapter is based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

@@

Unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.

>>>>

% ----------------------------------------------------------------------
\chapter{Improving Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc
% ----------------------------------------------------------------------

\paragraph{Abstract}

@@

More precisely, on one hand, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time. If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
%
On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real radio hardware.

On the other hand, we present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
Interestingly, the simpler heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.

\vfill{}

@@

This Chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

@@

We already explained in Chapter~\ref{chapter:1} that
unlicensed bands are more and more used and considered for mobile and LAN communication standards (WiFi, LTE-U), and for Internet of Things (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands will cause performance drop, and could even compromise IoT promises.

### 2019-02-27 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Proofreading of Chapter 4. No new science since the papers... but it's cleanly presented and unified.
Finishing Chapter:4

% For \rhoRand, \cite{Anandkumar11} essentially proved on the one hand that conditionnaly to a certain ``good'' event
% (all players have a correct ordering of all the $M$ best arms)
% proved to happen with high probability,
% the total number of collisions is upper bounded asymptotically by ${2 M - 1 \choose M}$,
% which is constant but grows very quickly as $M$ grows.
% And on the other hand, they proved that the total number of steps when this event is violated is $\bigO{\log T}$, with large constants in the asymptotic notation.

@@

For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
that we can then bound by the number of changes of arms
(\cf{} proof in Appendix~\ref{proof:5:collisionsMCTopM}).
%
An interesting consequence of the proof of this result is that
it also bounds the number of \emph{switches of arms}, $\sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$,
and this additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
Even though minimizing switching was not a goal\footnote{Introducing \emph{switching costs}, like it was done in previous works, \eg, \cite{Koren17}, could be an interesting future work.},
this guarantee is interesting for Cognitive Radio applications,
where switching arms means reconfiguring a radio hardware, an operation that costs energy.
%and thus an algorithm guaranting a small number of switches is interesting.

>>>>

% For \rhoRand, \cite{Anandkumar11} essentially proved on one hand that conditionnaly to a certain ``good'' event
% (all players have a correct ordering of all the $M$ best arms)
% proved to happen with high probability,
% the total number of collisions is upper bounded asymptotically by ${2 M - 1 \choose M}$,
% which is constant but grows very quickly as $M$ grows.
% And on the other hand, they proved that the total number of steps when this event is violated is $\bigO{\log T}$, with large constants in the asymptotic notation.

@@

For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
that we can then bound by the number of changes of arms
(\cf{} proof in Appendix~\ref{proof:5:collisionsMCTopM}).
%
An interesting consequence of the proof of this result is that
it also bounds the number of \emph{switches of arms}, $\sum_{t=1}^T \Pr(A^j(t+1) \neq A^j(t))$,
and this additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
Even though minimizing switching was not a goal\footnote{Introducing \emph{switching costs}, like it was done in previous works, \eg, \cite{Koren17}, could be an interesting future work.},
this guarantee is interesting for Cognitive Radio applications,
where switching arms means re-configuring a radio hardware, an operation that costs energy.
%and thus an algorithm guaranting a small number of switches is interesting.
